Police Shootings and Racism in America

### 2020-11-09 police_eda.Rmd
results

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon))

@@

```{r merge-data}
final_df <- merge(all, census, by.x = "city", by.y = "city_name")
final_df <- merge(final_df, gun_provisions, by = "state")
final_df <- final_df %>% 
  select(-c(department, year))
final_df$NumberOfSubjects[which(final_df$NumberOfSubjects == 0)] <- 1
```

>>>>

numofficerna <- which(is.na(all$NumberOfOfficers))
replacement <- str_count(all$OfficerRace[numofficerna], ";")
replacement <- replacement + 1
all$NumberOfOfficers[numofficerna] <- replacement

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         omult = as.numeric(all$NumberOfOfficers != "1"),
         other = as.numeric(weapon == "other"), 
         replica = as.numeric(weapon == "replica"), 
         knife = as.numeric(weapon == "knife"), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon))

@@

```{r merge-data}
final_df <- merge(all, census, by.x = "city", by.y = "city_name")
final_df <- merge(final_df, gun_provisions, by = "state")
final_df <- final_df %>% 
  select(-c(department, year))
final_df$NumberOfSubjects[which(final_df$NumberOfSubjects == 0)] <- 1
final_df$NumberOfOfficers[which(final_df$NumberOfOfficers == "U")] <- 1
```

### 2020-11-09 police_modeling.Rmd
results

## Modeling
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lmtest)
```

@@

```{r}
fit1 <- glm(formula = fatal ~ SubjectRace, family = binomial, data=single_victims)

fit2 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013), family = binomial, data=single_victims)
lrtest(fit2, fit1)
summary(fit2)

# Victim Gender doesn't appear to be significant
fit3 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + SubjectGender, family = binomial, data=single_victims)
lrtest(fit3, fit2)

fit4 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob, family = binomial, data=single_victims)
lrtest(fit3, fit4)

fit5 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow, family = binomial, data=single_victims)
lrtest(fit5, fit4)

fit6 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol, family = binomial, data=single_victims)
lrtest(fit5, fit6)

fit7 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
lrtest(fit6, fit7)

fit8 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
lrtest(fit8, fit7)

fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + gun, family = binomial, data=single_victims)
lrtest(fit8, fit9)
summary(fit9)
```

```{r}
mult1 <- glm(formula = fatal ~ sb + sw + sl + I(black/city_pop_2013) + ob + ow + ol + ow:sb + gun, family = binomial, data=mult_victims)
summary(mult1)
```

>>>>

## Modeling
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lmtest)
library(lme4)
library(caret)
library(groupdata2) # fold()
```

@@

### Old Validation Code

```{r}
crossvalidate <- function(data, k, model, dependent, random = FALSE){
  # 'data' is the training set with the ".folds" column
  # 'k' is the number of folds we have
  # 'model' is a string describing a linear regression model formula
  # 'dependent' is a string with the name of the score column we want to predict
  # 'random' is a logical; do we have random effects in the model?

  # Initialize empty list for recording performances
  performances <- c()

  # One iteration per fold
  for (fold in 1:k){

    # Create training set for this iteration
    # Subset all the datapoints where .folds does not match the current fold
    training_set <- data[data$.folds != fold,]

    # Create test set for this iteration
    # Subset all the datapoints where .folds matches the current fold
    testing_set <- data[data$.folds == fold,]

    ## Train model

    # If there is a random effect,
    # use glmer() to train model
    # else use glm()

    if (isTRUE(random)){
      # Train linear mixed effects model on training set
      m <- glmer(model, training_set, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
    } else {
      # Train linear model on training set
      m <- glm(model, training_set, family = "binomial", control=glm.control(maxit=50))
    }

    ## Test model

    # Predict the dependent variable in the testing_set with the trained model
    predicted <- predict(m, testing_set, type="response", allow.new.levels = TRUE)
    fit.convert <- ifelse(predicted < 0.5, 0, 1)
    table(fit.convert, testing_set[[dependent]])
    performances[fold] <- mean(fit.convert == testing_set[[dependent]])
  }

  return(performances)
}
```

```{r}
getAccuracy <- function(v) {
  if (length(v) == 4) {
      return(((v[1] + v[4])/sum(v)) * 100)
  }
  if (length(v) == 2) {
    return((v[1]/sum(v)) * 100)
  }
}
```

## Single Victims

### Backwards Selection
```{r backwards-selection}
fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=single_victims)
step(fullmodel, direction = "backward")
```

Backwards selection gives: glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
    data = single_victims)

```{r backwards-selection}
fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train + city, family = binomial, data=single_victims)
step(fullmodel, direction = "backward")
```

When I put city in, it turns out to be in the backwards model too -- should probably do random effects with city. Formula: glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city, family = binomial, data = single_victims)

### Trying the models

```{r backwards-fit}
fitback <- glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
    data = single_victims)
summary(fitback)
```

```{r validate fitback}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_fitback <- train(as.factor(fatal) ~ SubjectRace + SubjectGender + ob + ow + 
    omult + gun + knife + unarmed + I(black/city_pop_2013),  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_fitback, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```

```{r validate fitback again}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + SubjectGender + ob + ow + omult + gun + knife + unarmed + I(black/city_pop_2013)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = FALSE)
p0
mean(p0)

back2 <- glm(m0, training, family = "binomial")
preds <- predict(rand_fit1, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r fit with city}
fitback1 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city, family = binomial, data = single_victims)
summary(fitback1)

back2 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:SubjectRace, family = binomial, data = single_victims)
lrtest(fitback1, back2)

back3 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:gun, family = binomial, data = single_victims)
lrtest(fitback1, back3)
```

Interactions between city and race lead to better fit, makes it seem like I should vary race by city perhaps

```{r validate fitback1}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_fitback1 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_fitback1, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```
```{r validate back2}
set.seed(1)
Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

mod_back2 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + city + city:SubjectRace,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
    )

pred <- predict(mod_back2, newdata=testing )
confusionMatrix(data=pred, as.factor(testing$fatal))
```

```{r random-effects}
rand_fit1 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (1 | city), data = single_victims, family = "binomial", nAGQ=50)
summary(rand_fit1)
coef(rand_fit1)
fixef(rand_fit1)
ranef(rand_fit1)

classDFrand_fit1 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit1),0))
rand_fit1pred <- xtabs(~ predicted + response, data = classDFrand_fit1)
getAccuracy(rand_fit1pred)
```

```{r validate rand_fit1}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit1 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit1, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r vary subject race by city}
rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit2)
coef(rand_fit2)
fixef(rand_fit2)
ranef(rand_fit2)

classDFrand_fit2 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit2),0))
rand_fit2pred <- xtabs(~ predicted + response, data = classDFrand_fit2)
getAccuracy(rand_fit2pred)
```

```{r validate rand_fit2}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (SubjectRace | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit2 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit2, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r vary subject race by city and make it vary intercept too}
rand_fit3 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
    replica + knife + unarmed + (1 + SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit3)
coef(rand_fit3)
fixef(rand_fit3)
ranef(rand_fit3)

classDFrand_fit3 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit3),0))
rand_fit3pred <- xtabs(~ predicted + response, data = classDFrand_fit3)
getAccuracy(rand_fit3pred)
```

```{r validate rand_fit3}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + SubjectRace | city)'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit3, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

```{r random effects 4}
rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit4)
coef(rand_fit4)
fixef(rand_fit4)
ranef(rand_fit4)

classDFrand_fit4 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit4),0))
rand_fit4pred <- xtabs(~ predicted + response, data = classDFrand_fit4)
getAccuracy(rand_fit4pred)
```

```{r validate rand_fit4}
set.seed(1)

Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
training <- single_victims[ Train, ]
testing <- single_victims[ -Train, ]

data <- fold(
  training,
  k = 5
) %>%
  arrange(.folds)

m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender'
p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
p0
mean(p0)

rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
preds <- predict(rand_fit3, testing, type = "response")
fit.convert <- ifelse(preds < 0.5, 0, 1)
table(fit.convert, testing[['fatal']])
mean(fit.convert == testing[['fatal']])
```

Single victims final model!!

## Multiple Victims
```{r backwards-selection}
fullmodel <- glm(fatal ~ sb  + sw + sl + so + sfem + smale + weapon_other + replica + knife + ob + ow + ol + oo + omale + ofem + omult + gun + unarmed + unknown_weapon +  I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=mult_victims)
step(fullmodel, direction = "backward")
```

### 2020-11-09 submissioniii.Rmd
results

### Data

@@

## II. Methods

@@

|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among all shootings with only one victim, it appears in \emph{Figure 2} that the victim is more often Black than some other race. However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally. Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small. This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.

@@

***Since A and O are so little in the single victims dataset, I will combine them.

### Model

|        Because I want to explore the relationship between the number of victims in a police shooting affects and whether the shooting is fatal or not, I decided to create two separate models: one for shootings with only one victim, and one for shootings with multiple victims. Both models are fixed effect logistic regression models because the range that the response variable can take in logistic regression is the most reasonable given that I am predicting whether a victim is shot fatally or not (labeled 1 and 0, respectively). From the Exploratory Data Analysis section, it is evident that victim race plays a significant role in whether the shooting is fatal or not. Thus, for both models, I started with a simple logistic regression model using only victim race as a predictor. From there, I began adding the other predictors that I wanted to explore one at a time and performed likelihood ratio tests with each new model iteration to make sure that the added variable improves the previous model. This means that the likelihood ratio test's p-value is less than 0.05.

|        The final model for single-victim shootings is written out below:
```{r, eval = F}
fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
```

\begin{center}
$$
log(\frac{p_i}{(1-p_i)}) = \beta_0 + \sum_{j=1}^{4} \beta_{j} I(SubjectRace_i = SubjectRace[j]) +
$$
$$
\sum_{j=5}^{8}  \beta_{j} black\_officer\_present_i * I(SubjectRace_i = SubjectRace[j-4]) + \beta_9\ (\frac{black_i}{city\_pop_i}) +
$$
$$
\beta_{10}\ black\_officer\_present_i + \beta_{11}\ white\_officer\_present_i + \beta_{12}\ latino\_officer\_present_i + \beta_{13} \ victim\_armed\_gun_i
$$
$$
SubjectRace = [\text{Black, Latino, Other, White}]
$$
\end{center}

@@

## III. Preliminary Results

TO DO: 
- fix my models -- backward selection 
- random effects by city, maybe race | city 
- combine other with asian? 
- more interactions

@@

@@

## References

@@

## Appendix

>>>>

## II. Data

@@

@@

|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among all shootings with only one victim, it appears in \emph{Figure 2} that the victim is more often Black than some other race. However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally. Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small. Because of this, we will combine Asian and Other race types into one single race category called "Other" for both officers and victims for our models.

@@

## III. Methods

|        Because I want to explore the relationship between the number of victims in a police shooting affects and whether the shooting is fatal or not, I decided to create two separate models: one for shootings with only one victim, and one for shootings with multiple victims. Both models are random effect logistic regression models because the range that the response variable can take in logistic regression is the most reasonable given that I am predicting whether a victim is shot fatally or not (labeled 1 and 0, respectively). Additionally, I will be using random effects to control for correlation within cities.

|        To create these two models, I first performed backwards selection using simple logistic regression models on all relevant predictors in order to get a sense of which predictors are most statistically significant in prediction shooting fatality. Then, I added potential interactions that I wanted to explore one at a time and performed likelihood ratio tests with each new model iteration to check if the added interaction improved the previous model. Whether an interaction was statistically significant or not was determined by whether the likelihood ratio test's p-value was less than 0.05. Finally, I put these variables obtained from backwards selection into a random effects linear regression model with city as the random intercept and evaluated the fit of this model using cross validation, which I will describe in more detail in the Results section. From there, I tweaked the model by experimenting with additional predictors, interactions, and random effects between city and other predictors in order to obtain the best fitting model.

|        The final model for single-victim shootings is written out below:
```{r, eval = F}
rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
```

\begin{center}
$$
log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ SubjectRace_i + \beta_2\ OfficerBlack_i + \beta_3\ OfficerWhite_i + \beta_4\ MultipleOfficers_i
$$
$$
+ \beta_{5j}\ VictimWithGun_{ij} + \beta_6\ VictimWithOther_i + \beta_7\ VictimWithReplica_i + \beta_8\ VictimWithKnife_i + \beta_9\ VictimUnarmed_i + b_i
$$
$$
b_i \sim N(0, \sigma_b^2)
$$
$$
SubjectRace = \text{{White*, Black, Latino, Other}}
$$
\end{center}
In the model above, it is important to note that the predictor for whether the victim was carrying a gun is varied by city. This means that there is both a random intercept for city as well as a random slope for whether the victim was carrying a gun.

@@

## IV. Results

### Single Victims Model

```{r random effects 4}
rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(rand_fit4)
coef(rand_fit4)
fixef(rand_fit4)
ranef(rand_fit4)
```

### Multiple Victims Model

### Model Validation and Diagnostics

### Sensitivity Analysis
TO DO: 
- fix my models -- backward selection 
- random effects by city, maybe race | city 
- more interactions

@@

## V. Discussion and Conclusion

@@

## VI. References

@@

## VII. Appendix
